### 2019-03-28 Manuscript/Manuscript.tex
checkpoint after initial implementation of changes in response to reviewer comments

Outlines of the original surface extent of LIPs through the Phanerozoic (Fig. \ref{fig:LIP_map}) were slightly modified from the compilation of \citet{Ernst2017a}, and emplacement ages were taken from the literature (Table \ref{tab:LIPs}). The compilation of \citet{Ernst2017a} seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs. These polygons encapsulate all of the rocks associated with a given LIP, including dikes, sills, and layered intrusions, in addition to subaerial volcanics (Fig. \ref{fig:LIP_map}). For some LIPs, this approach may lead to an overestimate of original surface extent, given that subsurface intrusions could extend over a broader area than the surface volcanics. The polygons also assume complete surface coverage between wide-spread remnants, creating further potential for these outlines being over-estimates. However, this approach likely provides the best estimates available of original extent for ancient LIPs. The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the construction of the original extent polygons by \citet{Ernst2017a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).

@@

To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2019a}; Fig. \ref{fig:LIP_Areas}E). The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state. The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals. Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states. We take two approaches for comparison between the LIP area reconstruction and the record of ice extent. The first is to analyze the correlation between LIP area and the extent of ice away from the pole (Fig. \ref{fig:LIP_correlation}). The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles). This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.

@@

To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig. \ref{fig:LIP_Areas}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig. \ref{fig:LIP_correlation}). This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system. For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics. With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected. Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.

@@

The results from this analysis indicate that when the entire LIP database is considered, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig. \ref{fig:LIP_correlation}). While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that LIP area is not the fundamental control on Earth's long-term climate state.

@@

\begin{table}[h!]
\scriptsize
\caption{Phanerozoic large igneous provinces (and the Franklin).}
\resizebox{\textwidth}{!}{
\begin{tabular}{lc>{\raggedright}p{4cm}cc>{\raggedright}p{4cm}ccl}
\toprule
\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original} & \textbf{present} & \textbf{present area ref.} & \textbf{present/} & \textbf{half-life}$^{1}$ & \textbf{buried?} \\
 & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
 & & & \textbf{(Mkm$^{2}$)} & \textbf{(Mkm$^{2}$)} & & & & \\
\midrule
Columbia River & 16 & \cite{Kasbohm2018a} & 0.68 & 0.38 & \cite{Buchan2004a} & 0.56 & 19.2 & no \\
 & & & & & & & & \\
Afar & 30 & \cite{Courtillot2003a} & 2.05 & 0.63 & \cite{Coffin2006a} & 0.31 & 17.7 & partial \\
 & & & & & & & & \\
NAIP & 62 & \cite{Larsen2015a} & 1.07 & 0.29 & \cite{Buchan2004a, Coffin2006a} & 0.27 & 33.0 & partial \\
 & & & & & & & & \\
Deccan & 66 & \cite{Schoene2014a} & 0.83 & 0.56 & \cite{Coffin2006a} & 0.68 & 116.6 & no \\
 & & & & & & & & \\
Seychelles & 66 & \cite{Schoene2014a} & 0.46 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
 & & & & & & & & \\
Madagascar & 90 & \cite{Cucciniello2010a} & 0.63 & 0.03 & \cite{Coffin2006a} & 0.05 & 20.8 & no \\
 & & & & & & & & \\
Caribbean-Colombian & 94 & \cite{Loewen2013a} & 0.71 & 0.13 & \cite{Coffin2006a} & 0.18 & 37.6 & no \\
 & & & & & & & & \\
HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Moosdorf2015a} & 0.04 & 20.7 & no \\
 & & & & & & & & \\
EQUAMP & 131 & \cite{Hollanda2016a} & 0.66 & 0.01 & \cite{Hollanda2016a} & 0.01 & 20.5 & no \\
 & & & & & & & & \\
Comei & 132 & \cite{Zhu2009a} & 0.11 & - & - & - & - & no \\
 & & & & & & & & \\
Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
 & & & & & & & & \\
Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018a} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
 & & & & & & & & \\
Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\
 & & & & & & & & \\
NW Australia Margin & 160 & \cite{Pirajno2012a} & 0.62 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
 & & & & & & & & \\
Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{2}$ & 0.05 & 41.3 & no \\
 & & & & & & & & \\
Ferrar & 183 & \cite{Burgess2015a} & 0.18 & - & - & - & - & no \\
 & & & & & & & & \\
CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{3}$ & 0.02 & 35.7 & partial \\
 & & & & & & & & \\
Siberia & 252 & \cite{Burgess2015b} & 3.46 & 0.47 & \cite{Coffin2006a} & 0.14 & 87.5 & no \\
 & & & & & & & & \\
Emeishan & 259 & \cite{Zhou2002a} & 0.71 & 0.06 & \cite{Coffin2006a} & 0.09 & 72.9 & no \\
 & & & & & & & & \\
Panjal-Qiangtang & 283 & \cite{Zhai2013a} & 0.11 & - & - & - & - & no \\
 & & & & & & & & \\
Tarim & 290 & \cite{Xu2014a} & 0.35 & - & - & - & - & no \\
 & & & & & & & & \\
Magdalen & 360 & \cite{Brendan_Murphy1999a} & 0.42 & - & - & - & - & no \\
 & & & & & & & & \\
Vilyui & 374 & \cite{Ricci2013a} & 1.14 & - & - & - & - & no \\
 & & & & & & & & \\
Kola-Dnieper & 380 & \cite{Arzamastsev2014a} & 5.90 & - & - & - & - & no \\
 & & & & & & & & \\
Suordakh & 450 & \cite{Khudoley2013a} & 0.02 & - & - & - & - & no \\
 & & & & & & & & \\
Kalkarindji & 511 & \cite{Jourdan2014a} & 3.54 & 0.17 & \cite{Thorne2014a} & 0.05 & 116.3 & no \\
 & & & & & & & & \\
Franklin & 720 & \cite{Denyszyn2009a} & 2.62 & 0.04 & \cite{Buchan2004a} & 0.02 & 121.8 & no \\
\bottomrule
\end{tabular}
}
\begin{tablenotes}[para,flushleft]
$^{1}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.

$^{2}$from ArcGIS compilation produced by M. de Kock for the LIPs Reconstruction Project \citep{Ernst2013a}.

$^{3}$from ArcGIS compilation produced by A. Marzoli and L. Parisio for the LIPs Reconstruction Project \citep{Ernst2013a}.
\end{tablenotes}
\label{tab:LIPs}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealp{Ernst2017a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
	\label{fig:LIP_map}
\end{center}
\end{figure}

>>>>

Outlines of the original surface extent of continental LIPs through the Phanerozoic (Fig. \ref{fig:LIP_map}) were slightly modified (to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons) from the compilation of \citet{Ernst2017a} and \citet{Ernst2019a}, and emplacement ages were taken from the literature (Table \ref{tab:LIPs}). The LIP original surface extent compilation seeks to reconstruct the original surface extent of LIPs with the caveat that there can be significant uncertainty with doing so, particularly for older more deeply eroded LIPs. These polygons encapsulate all of the rocks associated with a given LIP, including dikes, sills, and layered intrusions, in addition to subaerial volcanics (Fig. \ref{fig:LIP_map}). For some LIPs, this approach may lead to an overestimate of original surface extent, given that subsurface intrusions could extend over a broader area than the surface volcanics. The polygons also assume complete surface coverage between wide-spread remnants, creating further potential for these outlines being over-estimates. However, this approach likely provides the best estimates available of original extent for ancient LIPs. The extents of presently-exposed volcanics associated with LIPs that were used for present-day area estimates, and the construction of the original extent polygons by \citet{Ernst2017a} and \citet{Ernst2019a}, were taken from a number of resources including the PLATES compilation \citep{Coffin2006a} and more recent compilation efforts associated with the LIPs Reconstruction Project (\citealp{Ernst2013a}; Table \ref{tab:LIPs}).

@@

To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic (\citealp{Macdonald2019a}; Fig. \ref{fig:LIP_Areas}E). The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state. The land ice record is an imperfect tracker of climate, as it is insensitive to changes in temperature during non-glacial intervals, is influenced by additional factors such as the physical geography of the continents during glacial intervals, and is potentially vulnerable to removal from the observable geologic record via erosion and/or burial. Nevertheless, it forms a physical record of Earth's climate through time and delineates glacial and non-glacial climate states. We take two approaches for comparison between the LIP area reconstruction and the record of ice extent. The first is to analyze the correlation between LIP area and the extent of ice away from the pole (Fig. \ref{fig:LIP_correlation}). The second is to consider the overlap between intervals of high LIP area (defined as LIP area \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles). This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.

@@

To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig. \ref{fig:LIP_Areas}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig. \ref{fig:LIP_correlation}). This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system. For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial interval simulations correlate better with the LIP area in the tropics. With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected. Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.

@@

The results from this analysis indicate that when the entire LIP database is considered, there is no significant relationship between total LIP area or LIP area in the tropics and the extent of continental ice sheets (Fig. \ref{fig:LIP_correlation}). While this result need not imply that there is no increase in global weatherability from the emplacement of LIPs, it does suggest that changes in planetary weatherability associated with LIPs are not the fundamental control on whether Earth is in a glacial or non-glacial climate state.

@@

\begin{table}[h!]
\scriptsize
\caption{Phanerozoic large igneous provinces (and the Franklin).}
\resizebox{\textwidth}{!}{
\begin{tabular}{lc>{\raggedright}p{4cm}cc>{\raggedright}p{4cm}ccl}
\toprule
\textbf{name} & \textbf{age} & \textbf{age ref.} & \textbf{original$^{1}$} & \textbf{present} & \textbf{present area ref.} & \textbf{present/$^{2}$} & \textbf{half-life}$^{3}$ & \textbf{buried?} \\
 & \textbf{(Ma)} & & \textbf{area} & \textbf{area} & & \textbf{original} & \textbf{(Myr)} & \\
 & & & \textbf{(Mkm$^{2}$)} & \textbf{(Mkm$^{2}$)} & & & & \\
\midrule
Columbia River & 16 & \cite{Kasbohm2018a} & 0.68 & 0.38 & \cite{Buchan2004a} & 0.56 & 19.2 & no \\
 & & & & & & & & \\
Afar & 30 & \cite{Courtillot2003a} & 2.05 & 0.63 & \cite{Coffin2006a} & 0.31 & 17.7 & partial \\
 & & & & & & & & \\
NAIP & 62 & \cite{Larsen2015a} & 1.07 & 0.29 & \cite{Buchan2004a, Coffin2006a} & 0.27 & 33.0 & partial \\
 & & & & & & & & \\
Deccan & 66 & \cite{Schoene2014a} & 0.83 & 0.56 & \cite{Coffin2006a} & 0.68 & 116.6 & no \\
 & & & & & & & & \\
Seychelles & 66 & \cite{Schoene2014a} & 0.46 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
 & & & & & & & & \\
Madagascar & 90 & \cite{Cucciniello2010a} & 0.63 & 0.03 & \cite{Coffin2006a} & 0.05 & 20.8 & no \\
 & & & & & & & & \\
Caribbean-Colombian & 94 & \cite{Loewen2013a} & 0.71 & 0.13 & \cite{Coffin2006a} & 0.18 & 37.6 & no \\
 & & & & & & & & \\
HALIP & 95 & \cite{Kingsbury2018a} & 3.60 & 0.15 & \cite{Moosdorf2015a} & 0.04 & 20.7 & no \\
 & & & & & & & & \\
EQUAMP & 131 & \cite{Hollanda2016a} & 0.66 & 0.01 & \cite{Hollanda2016a} & 0.01 & 20.5 & no \\
 & & & & & & & & \\
Comei & 132 & \cite{Zhu2009a} & 0.11 & - & - & - & - & no \\
 & & & & & & & & \\
Bunbury & 132 & \cite{Zhu2009a} & 0.03 & 0.00 & \cite{Thorne2014a} & 0.05 & 30.9 & no \\
 & & & & & & & & \\
Parana-Etendeka & 135 & \cite{Florisbal2014a, Almeida2018a} & 3.12 & 0.40 & \cite{Coffin2006a} & 0.13 & 45.7 & partial \\
 & & & & & & & & \\
Trap & 140 & \cite{Ernst2001a} & 0.03 & 0.00 & \cite{Ernst2001a} & 0.00 & 0.0 & no \\
 & & & & & & & & \\
NW Australia Margin & 160 & \cite{Pirajno2012a} & 0.62 & 0.00 & \cite{Coffin2006a} & 0.00 & 0.0 & yes \\
 & & & & & & & & \\
Karoo & 183 & \cite{Burgess2015a} & 3.21 & 0.15 & de Kock compilation$^{4}$ & 0.05 & 41.3 & no \\
 & & & & & & & & \\
Ferrar & 183 & \cite{Burgess2015a} & 0.18 & - & - & - & - & no \\
 & & & & & & & & \\
CAMP & 201 & \cite{Blackburn2013a} & 11.46 & 0.23 & Marzoli and Parisio compilation$^{5}$ & 0.02 & 35.7 & partial \\
 & & & & & & & & \\
Siberia & 252 & \cite{Burgess2015b} & 3.46 & 0.47 & \cite{Coffin2006a} & 0.14 & 87.5 & no \\
 & & & & & & & & \\
Emeishan & 259 & \cite{Zhou2002a} & 0.71 & 0.06 & \cite{Coffin2006a} & 0.09 & 72.9 & no \\
 & & & & & & & & \\
Panjal-Qiangtang & 283 & \cite{Zhai2013a} & 0.11 & - & - & - & - & no \\
 & & & & & & & & \\
Tarim & 290 & \cite{Xu2014a} & 0.35 & - & - & - & - & no \\
 & & & & & & & & \\
Magdalen & 360 & \cite{Brendan_Murphy1999a} & 0.42 & - & - & - & - & no \\
 & & & & & & & & \\
Vilyui & 374 & \cite{Ricci2013a} & 1.14 & - & - & - & - & no \\
 & & & & & & & & \\
Kola-Dnieper & 380 & \cite{Arzamastsev2014a} & 5.90 & - & - & - & - & no \\
 & & & & & & & & \\
Suordakh & 450 & \cite{Khudoley2013a} & 0.02 & - & - & - & - & no \\
 & & & & & & & & \\
Kalkarindji & 511 & \cite{Jourdan2014a} & 3.54 & 0.17 & \cite{Thorne2014a} & 0.05 & 116.3 & no \\
 & & & & & & & & \\
Franklin & 720 & \cite{Denyszyn2009a} & 2.62 & 0.04 & \cite{Buchan2004a} & 0.02 & 121.8 & no \\
\bottomrule
\end{tabular}
}
\begin{tablenotes}[para,flushleft]
$^{1}$calculated from the continental portion of polygons within the LIP original surface extent compilation of \citet{Ernst2017a} and \citet{Ernst2019a}.

$^{2}$present area / original area

$^{3}$assuming exponential decay with form $N(t) = 2^{t/t_{1/2}}$.

$^{4}$from ArcGIS compilation produced by M. de Kock for the LIPs Reconstruction Project \citep{Ernst2013a}.

$^{5}$from ArcGIS compilation produced by A. Marzoli and L. Parisio for the LIPs Reconstruction Project \citep{Ernst2013a}.
\end{tablenotes}
\label{tab:LIPs}
\end{table}

\begin{table}[h!]
\scriptsize
\caption{Statistics of correlation between large igneous province area and ice extent.}
\resizebox{\textwidth}{!}{
\begin{tabular}{>{\raggedright}p{2.5cm}cccccccccccccccc}
\toprule
\textbf{} & \multicolumn{4}{c}{\textbf{total}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$10$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$15$^{\circ}$}} & \multicolumn{4}{c}{\textbf{within tropics $\pm$20$^{\circ}$}} \\
\textbf{} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} & \multicolumn{2}{c}{\textbf{correlation}} & \multicolumn{2}{c}{\textbf{\% overlap}} \\
\textbf{scenario} & \textbf{val.}$^{1}$ & \textbf{p-val.}$^{2}$ & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} & \textbf{val.} & \textbf{p-val.} \\
\midrule
t$_{1/2}$ = 36~Myr & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
&&&&&&&&&&&&&&&& \\
t$_{1/2}$ = 36~Myr + 50\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
&&&&&&&&&&&&&&&& \\
t$_{1/2}$ = 36~Myr + 100\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
&&&&&&&&&&&&&&&& \\
t$_{1/2}$ = 120~Myr + 100\% burial & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X & X \\
\bottomrule
\end{tabular}
}
\begin{tablenotes}[para,flushleft]
$^{1}$`val.' refers to the computed correlation coefficient/\% overlap between LIP area and the actual ice extent record.

$^{2}$`p-val.' refers to the fraction of randomly timed glacial interval simulations that correlate/overlap better with LIP area than the actual ice extent record (i.e. the p-value with respect to the null hypothesis of no correlation/overlap).
\end{tablenotes}
\label{tab:stats}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=\textwidth]{Manuscript/Figures/LIP_Map.pdf}
	\caption{Map of current extent of volcanic lithologies associated with LIPs that erupted between 520~Ma and the present, as well as the estimates of the initial LIP area used in the area analysis (modified slightly from \citealt{Ernst2017a} and \citealt{Ernst2019a} to ensure that all currently exposed volcanic lithologies are encapsulated by the initial LIP area polygons).}
	\label{fig:LIP_map}
\end{center}
\end{figure}
